\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2021-08-02 manuscript.Rnw
Minor edits to prose and references

%% -- Discussion ---------------------------------------------------------------
\section{Discussion} \label{sec:discussion}
This paper formally describes the economic models and cost-effectiveness framework used by the \pkg{hesim} package. An illustrative example was provided for a three state model commonly used to evaluate treatments in oncology. However, since \pkg{hesim} aims to be flexible, the example only covers a subset of the relevant modeling approaches and functionality. Additional examples as well as documentation for all functions and classes are provided on the package website (\url{https://hesim-dev.github.io/hesim/}).

@@

A design decision that is worth commenting on is the use of \pkg{R6}. While it may not be ``traditional'' \proglang{R}, we believe it fits our modular design nicely. It is consistent with our conceptualization of the economic modeling process (Figure~\ref{fig:econ-eval-process-hesim}) in that we can have separate submodels for disease progression, costs, and utility, which can be combined into an economic model. A particularly important benefit of \pkg{R6} is that methods can belong to objects. In our framework, this means that submodels can not only contain inputs for the economic model, but can also generate outputs. For example, the disease model of a PSM not only contains the parameters of the survival models and the input data required for prediction, but can also generate relevant outcomes such as survival curves or hazard functions. Likewise, the disease model of an iCTSTM not only stores parameters and input data, but can be used to simulate trajectories through a multi-state model and compute state probabilities. This makes it easier to check that the submodels produce desired outputs, compare submodels with different underlying modeling assumptions (e.g., different parametric survival distributions), and to use the package in a more modular way. The latter may be especially relevant to users who only want to use some features of the package; for, instance a user interested only in epidemiological outcomes could use the disease progression submodel without needing to use the utility or cost submodels. All that said, we utilize \code{S3} whenever it makes sense as it is more familiar to most \proglang{R} users.

@@

There are of course features in other \proglang{R} packages that are not in the current version of \pkg{hesim}. For example, \pkg{heemod} has an explicit function, \code{calibrate_model()}, for calibrating model parameters that are difficult to measure. Furthermore, tunnel states can be automatically created (rather than needing to be created manually) with \pkg{heemod} using the \code{state_time} variable so that costs or transition probabilities can very as a function of time in a health state. \pkg{hesim} is also limited in its support for value of information analyses. For instance, neither the expected value of partial perfect information (EVPPI) or the expected value of sample information (EVSI) can be computed with \pkg{hesim}, whereas the EVPPI can be computed with \pkg{BCEA}, \pkg{dampack}, and \pkg{SAVI} and EVSI can be computed with \pkg{EVSI} and \pkg{dampack}. In this respect, these packages are complementary and simulation output from \pkg{hesim} could be used as inputs to their more advanced value of information functions.

A number of extensions of the current version of \pkg{hesim} would be helpful. First, we plan to extend the individual-level simulation so that users can update input data during the simulation as a function of time elapsed and/or the health states transitioned to. This would allow potentially time-varying variables such as age to update each time a patient transitions to a new state. Second, we would like to add more examples to ensure that \pkg{hesim} is fit for purpose for a wide range of applications in health economics. Third, more complex models of utility \citep{kharroubi2007modelling, hernandez2013relationship} and costs \citep{nixon2005methods} could help better capture heterogeneity in preferences and spending. Fourth, it would be useful to add functionality that automatically creates tunnel states for users in cDTSTMs. Fifth, equity considerations could be incorporated by adding support for distributional CEA \citep{asaria2015distributional, asaria2016distributional, cookson2017using}. Sixth, functions could be added to convert \pkg{hesim} simulation output into formats required by other packages that can analyze the outputs of health economics models, such as those with more advanced value of information analysis capabilities. Finally, a larger aspiration is to provide support for parameterization via NMA, particularly for estimation of relative treatment effects in survival (e.g., for PSMs) and multi-state models. However, since existing \proglang{R} packages are lacking in this area, this could involve creation of a separate NMA package and a new dependency for \pkg{hesim}.

>>>>

%% -- Discussion ---------------------------------------------------------------
\section{Discussion} \label{sec:discussion}
This paper formally describes the economic modeling and cost-effectiveness framework used by the \pkg{hesim} package. An illustrative example was provided for a three state model commonly used to evaluate treatments in oncology. However, since \pkg{hesim} aims to be flexible, the example only covers a subset of the relevant modeling approaches and functionality. Additional examples as well as documentation for all functions and classes are provided on the package website (\url{https://hesim-dev.github.io/hesim/}).

@@

A design decision that is worth commenting on is the use of \pkg{R6}. While it may not be ``traditional'' \proglang{R}, we believe it fits our modular design nicely. It is consistent with our conceptualization of the economic modeling process (Figure~\ref{fig:econ-eval-process-hesim}) in that it is natural to have classes representing separate submodels for disease progression, costs, and utility, which can be combined into a single class representing an economic model. A particularly important benefit of \pkg{R6} is that methods can belong to objects. In our framework, this means that submodels can not only contain inputs for the economic model, but can also generate outputs. For example, the disease model of a PSM not only contains the parameters of the survival models and the input data required for prediction, but can also generate relevant outcomes such as survival curves or hazard functions. Likewise, the disease model of an iCTSTM not only stores parameters and input data, but can be used to simulate trajectories through a multi-state model and compute state probabilities. This makes it easier to check that the submodels produce desired outputs, compare submodels with different underlying modeling assumptions (e.g., different parametric survival distributions), and to use the package in a more modular way. The latter may be especially relevant to users who only want to use some features of the package; for, instance a user interested only in epidemiological outcomes could use the disease progression submodel without needing to use the utility or cost submodels. All that said, we utilize \code{S3} whenever it makes sense as it is more familiar to most \proglang{R} users.

@@

There are of course features in other \proglang{R} packages that are not in the current version of \pkg{hesim}. For example, \pkg{heemod} has an explicit function, \code{calibrate_model()}, for calibrating model parameters that are difficult to measure. Furthermore, tunnel states can be automatically created (rather than needing to be created manually) with \pkg{heemod} using the \code{state_time} variable so that costs or transition probabilities can very as a function of time in a health state. \pkg{hesim} is also limited in its support for value of information analyses. In particular, neither the expected value of partial perfect information (EVPPI) or the expected value of sample information (EVSI) can be computed with \pkg{hesim}, whereas the EVPPI can be computed with \pkg{BCEA}, \pkg{dampack}, and \pkg{SAVI} and EVSI can be computed with \pkg{EVSI} and \pkg{dampack}. In this respect, these packages are complementary and simulation output from \pkg{hesim} could be used as inputs to their more advanced value of information functions.

A number of extensions of the current version of \pkg{hesim} would be helpful. First, we plan to extend the individual-level simulation so that users can update input data during the simulation as a function of time elapsed and/or the health states transitioned to. This would allow potentially time-varying variables such as age to update each time a patient transitions to a new state. Second, we would like to add more examples to ensure that \pkg{hesim} is fit for purpose for a wide range of applications in health economics. Third, more complex models of utility \citep{kharroubi2007modelling, hernandez2013relationship} and costs \citep{nixon2005methods} could help better capture heterogeneity in preferences and spending. Fourth, it would be useful to add functionality that automatically creates tunnel states for users in cDTSTMs. Fifth, equity considerations could be incorporated by adding support for distributional CEA \citep{asaria2015distributional, asaria2016distributional, cookson2017using}. Sixth, it would be convenient if \pkg{hesim} simulation output could be converted into formats required by other packages that can analyze the outputs of health economics models, such as those with more advanced value of information analysis capabilities. Finally, a larger aspiration is to provide support for parameterization via NMA, particularly for estimation of relative treatment effects in survival (e.g., for PSMs) and multi-state models. However, since existing \proglang{R} packages are lacking in this area, this could involve creation of a separate NMA package and a new dependency for \pkg{hesim}.
